---
output: 
  pdf_document:
    template: preprint_template.tex
    fig_caption: true
    keep_tex: false
    citation_package: natbib
  word_document: 
    reference_docx: doc_template.docx
title: "Intraspecific Variation Amplifies Spatial and Temporal Patterns in Vulnerability to Warming"
author:
  - name: Matthew Sasaki
    affiliation: University of Massachusetts Lowell
    location: 
  - name: Melissa Pespeni
    affiliation: University of Vermont
    location: 
abstract: "ABSTRACT HERE."
keywords: "keyword1, keyword2, keyword3"
contact: "**Correspondance**: Matthew_Sasaki@uml.edu" 
current: "This preprint is current as of `r format(Sys.time(), '%B %d, %Y')`"
repo: "**Code**: ___"
bibliography: references.bib
biblio-style: bib-style
geometry: margin=1in
fontsize: 10pt
graphics: yes
header-includes: 
  \usepackage{caption}    
  \captionsetup[figure]{font=small, margin=20pt, justification=centering}
  \captionsetup[table]{font=small, margin=20pt, justification=centering}
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r echo = F}
if(knitr::is_latex_output()){
  knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  message = FALSE,
  warning = FALSE,
  collapse = T)
}
```

# Introduction

# Methods

## Field Collections and Phenotypic Measurements

## Warming Tolerance

## Library Prep and Sequencing

## Clade Assignment

## Statistical Analysis

linear mixed effects model: CTmax ~ body size, temperature, and clade with random effect of tube and a random effect of site along with random slope for collection temperature

Marginal mean and temperature trend for each clade 

# Results

## Environmental Data

Figure 1: Map of sites (A) and collection temperatures (B)

Supp. Fig. 1: Continuous temperature profiles for the regions

Supp. Fig. 2: Collection temperature seasonality against latitude (A) and maximum collection temperature (B)

## COI Clade Distributions
Figure X: Clade plot (summarizes the clade x season x site patterns; just the 'high confidence' estimates - No hudsonica and only tonsa individuals where at least 85% of the reads matched to a single clade)

Supp. Fig. X: Individual clade prop plot (shows the proportion of reads for each individual that matched to the different clades)

## Seasonal phenotypic patterns
Supp. Fig. X: Temperature ramping records

Figure X: CTmax vs. season (A) and size vs. season (B) for each site

Supp. Fig. X: CTmax vs. season for each site, with collection temperature included (ctmax-ind-pops-season)


## Clade-specific patterns

Figure X: CTmax and body size vs. collection temperature, separated by clade

Supp. Fig. X: CTmax vs. collection temperature, separated by clade and faceted for each site to show the concordance in the pattern across locations

Figure X: Clade marginal means (i.e. genetic variation in thermal limit) and the marginal trend (i.e. combined effect of local adaptation and plasticity)

Figure X: Clade proportions in each sample vs. collection temp.

Supp. Fig. X: Contour plot, faceted by clade, showing the scaled density of each clade across the temperature x salinity space. Individual occurences across this space are included as jittered points to show the distribution and abundance of records. 

# Discussion 

## Study Summary 

## Space-for-time 

## Relevance of intraspecific variation (plasticity, local adaptation, cryptic variation like clades)



```{r, figure, echo = F}
#| fig.cap = "CAPTION HERE."

plot(1)
#knitr::include_graphics("../Output/Figures/_PATH_") #Use this when filling in figures from analysis 
```

```{r, table, echo = F}
knitr::kable(x = 1, #Replace 1 with whatever object you want included as a table
      digits = 3, align = "c",
      caption = "CAPTION HERE")
```

# Discussion

\newpage

```{=tex}
\beginsupplement
```
# Supplementary Material

```{r supp-fig-1, echo = F}
#| fig.cap = "CAPTION HERE."

plot(1)
#knitr::include_graphics("../Output/Figures/_PATH_") #Use this when filling in figures from analysis 
```
