---
title: "Comparing seasonal and latitudinal patterns in thermal adaptation"
date: "`r Sys.Date()`"
output: 
  html_document:
          code_folding: hide
          code_download: true
          toc: true
          toc_float: true
  github_document:
          html_preview: false
          toc: true
          toc_depth: 3
---

```{r setup, include=T, message = F, warning = F, echo = F}

knitr::opts_chunk$set(
  echo = knitr::is_html_output(),
  fig.align = "center",
  fig.path = "../Figures/markdown/",
  dev = c("png", "pdf"),
  message = FALSE,
  warning = FALSE,
  collapse = T
)

theme_matt = function(base_size = 18,
                      dark_text = "grey20"){
  mid_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[2]
  light_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[3]
  
  theme_pubr(base_family="sans") %+replace% 
    theme(
      panel.background  = element_rect(fill="transparent", colour=NA), 
      plot.background = element_rect(fill="transparent", colour=NA), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      text = element_text(colour = mid_text, lineheight = 1.1),
      title = element_text(size = base_size * 1.5,
                           colour = dark_text),
      axis.text = element_text(size = base_size,
                               colour = mid_text),
      axis.title.x = element_text(size = base_size * 1.2,
                                  margin = unit(c(3, 0, 0, 0), "mm")),
      axis.title.y = element_text(size = base_size * 1.2,
                                  margin = unit(c(0, 5, 0, 0), "mm"), 
                                  angle = 90),
      legend.text = element_text(size=base_size * 0.9),
      legend.title = element_text(size = base_size * 0.9, 
                                  face = "bold"),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25,"cm")
    )
}

theme_matt_facets = function(base_size = 18,
                             dark_text = "grey20"){
  mid_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[2]
  light_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[3]
  
  theme_bw(base_family="sans") %+replace% 
    theme(
      panel.grid = element_blank(),
      panel.background  = element_rect(fill="transparent", colour=NA), 
      plot.background = element_rect(fill="transparent", colour=NA), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      text = element_text(colour = mid_text, lineheight = 1.1),
      strip.text.x = element_text(size = base_size),
      title = element_text(size = base_size * 1.5,
                           colour = dark_text),
      axis.text = element_text(size = base_size,
                               colour = mid_text),
      axis.title.x = element_text(size = base_size * 1.2,
                                  margin = unit(c(3, 0, 0, 0), "mm")),
      axis.title.y = element_text(size = base_size * 1.2,
                                  margin = unit(c(0, 5, 0, 0), "mm"), 
                                  angle = 90),
      legend.text = element_text(size=base_size * 0.9),
      legend.title = element_text(size = base_size * 0.9, 
                                  face = "bold"),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25,"cm")
    )
}

site_cols = c("Key Largo" = "indianred4", 
              "Manatee River" = "coral", 
              "Ft. Hamer" = "coral3",
              "Tyler Cove" = "goldenrod1",
              "Ganey's Wharf" = "darkgoldenrod3", 
              "Esker Point" = "darkseagreen3",
              "Sawyer Park" = "palegreen4", 
              "St. Thomas de Kent Wharf" = "steelblue2",
              "Ritchie Wharf" = "steelblue4")

pop_cols = c("KL" = "indianred4", 
              "MR" = "coral", 
              "FH" = "coral3",
              "MD" = "goldenrod1",
              "GW" = "darkgoldenrod3", 
              "CT" = "darkseagreen3",
              "ME" = "palegreen4", 
              "TK" = "steelblue2",
              "RW" = "steelblue4")
```

## Main Message

By assaying thermal limits several times per year in natural populations distributed across a large latitudinal gradient, we highlight the importance of intra-specific variation for predicting vulnerability to climate change. Indeed, across the entire spatiotemporal framework examined here, we observed an ~12°C range in thermal limits. Both spatial and temporal variation in thermal limits is essential for robust predictions.


## Site Characteristics

```{r}
site_temps = full_data %>% 
  dplyr::select(site, lat, season, doy, collection_temp, collection_salinity) %>%  
  distinct() %>% 
  filter(doy > 100) 
```

Copepods were collected by surface tow from sites across the Western Atlantic at several times throughout the year. The sites are shown below along with surface temperatures and salinity were measured at the time of collection. Across the entire set of collections, temperature ranged from `r min(site_temps$collection_temp)`°C to `r max(site_temps$collection_temp)`°C.

```{r site-chars, fig.width=10, fig.height=6}
coords = site_data %>%
  dplyr::select(site, long, lat) %>%
  distinct()

site_map = map_data("world") %>% 
  filter(region %in% c("USA", "Canada")) %>% 
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group),
               fill = "lightgrey") + 
  coord_map(xlim = c(-85,-60),
            ylim = c(25, 48)) + 
  geom_point(data = coords,
             mapping = aes(x = long, y = lat, colour = site),
             size = 3) +
  scale_colour_manual(values = site_cols) + 
  labs(x = "Longitude", 
       y = "Latitude") + 
  theme_matt(base_size = 16)

site_temp_plot = ggplot(site_temps, aes(x = doy, y = collection_temp, colour = site)) + 
  geom_line(linewidth = 2) + 
  geom_point(size = 5) +
  scale_colour_manual(values = site_cols) + 
  labs(y = "Temperature (°C)",
       x = "Day of the Year") +
  theme_matt() + 
  theme(legend.position = "right")

ggarrange(site_map, site_temp_plot, common.legend = T, legend = "bottom")
```

Collections aimed to obtain copepods near the onset of peak temperatures, after peak temperatures, and then after temperatures cooled during the Fall. To put the temperature data we measured into this larger seasonal context, we examined continuous temperature records for each site. Continuous data is not available for all sites, however, so here we've pieced together daily temperature values from either local temperature sensors (sites in Florida and the Chesapeake) and high resolution satellite temperature data (Connecticut, Maine, and the Canadian sites). This satellite data comes from the NOAA 1/4° Daily Optimum Interpolation Sea Surface Temperature (OISST). 

These temperature profiles are shown below, with the temperatures measured during the time of collection included for comparison While collection temperatures in Maine did not always match the recorded daily averages, across all sites the temperature records do give a general sense of the timing of seasonal maxima (and the proximity of collections to the seasonal maximum temperature). In general, the first sample from each site fell just after the site reached the warmest period. The exception to that pattern is in Florida, where collections occurred after an extended period of high temperatures. 

```{r continuous-temps, fig.width=8, fig.height=6}
temp_profiles = temp_profiles %>% 
  filter(doy > 100) %>% 
  mutate(region = fct_relevel(region, "Florida", "Chesapeake", "Connecticut",
                              "Maine", "Shediac", "Miramichi"))

site_temps2 = site_temps %>% 
  mutate(region = case_when(
    site == "Manatee River" ~ "Florida",
    site == "Ft. Hamer" ~ "Florida",
    site == "Tyler Cove" ~ "Chesapeake",
    site == "Ganey's Wharf" ~ "Chesapeake",
    site == "Esker Point" ~ "Connecticut",
    site == "Sawyer Park" ~ "Maine",
    site == "St. Thomas de Kent Wharf" ~ "Shediac",
    site == "Ritchie Wharf" ~ "Miramichi"),
    region = fct_relevel(region, "Florida", "Chesapeake", "Connecticut",
                         "Maine", "Shediac", "Miramichi"))

ggplot(temp_profiles, aes(x = doy, y = temp_c)) + 
  facet_wrap(region~.) + 
  geom_point(data = site_temps2,
             aes(x = doy, y = collection_temp, colour = site),
             size = 3) +
  geom_line() + 
  scale_colour_manual(values = site_cols) + 
  labs(x = "Day of the Year", 
       y = "Mean Daily Temp. (°C)") + 
  theme_matt_facets() + 
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 320, hjust = 0, vjust = 0.5))
```


Exact locations for the sites are provided here: 

```{r site-table}
site_data %>%  
  arrange(lat) %>%  
  select("Site" = site, "Region" = region, "Lat" = lat, "Long" = long) %>% 
  knitr::kable(align = "c")
```

Nested within each of the three regions (South, Central, and Northern regions) are pairs of low and high salinity sites:    

```{r sal-table}
data.frame("Region" = c("South", "Central", "North"),
           "Low Salinity" = c("Ft. Hamer", "Ganey's Wharf", "Ritchie Wharf"),
           "High Salinity" = c("Manatee River", "Tyler Cove", "St. Thomas de Kent Wharf")) %>% 
  knitr::kable(align = "c")
```

\ 

There are fairly well-established divergences between high salinity and low salinity populations of *Acartia tonsa*. These sets of geographically proximate sites potentially provide independent comparisons of the effects of seasonality across salinity gradients. Shown here are the collection conditions for these pairs of sites. Temperature was typically similar across the pairs within each collection, while salinity differences were fairly stable across collections. 

```{r season-sal-comps, fig.width=8, fig.height=8}
season_cols = c("early" = "grey75", 
                "peak" = "grey50", 
                "late" = "grey25")

sal_regions = data.frame(region = rep(c("South", "Central", "North"), each = 2), 
                         site = c("Ft. Hamer", "Manatee River", 
                                  "Ganey's Wharf", "Tyler Cove", 
                                  "Ritchie Wharf", "St. Thomas de Kent Wharf"),
                         salinity = c("low", "high"))

sal_comps = full_data %>% 
  filter(site %in% sal_regions$site) %>% 
  inner_join(sal_regions, by = c("site")) %>% 
  select( region = region.y, site, salinity, season, doy, collection_temp, collection_salinity,
          size, ctmax, warming_tol) %>% 
  mutate(salinity = fct_relevel(salinity, "low", "high"),
         region = fct_relevel(region, "South", "Central", "North"))

sal_comp_temps = sal_comps %>%  
  select(salinity, season, region, collection_temp, collection_salinity) %>% 
  distinct() %>% 
  ggplot(aes(x = salinity, y = collection_temp, colour = season, group = season)) + 
  facet_wrap(region~.) + 
  geom_line(linewidth = 1.5) + 
  geom_point(size = 4) + 
  scale_colour_manual(values = season_cols) + 
  labs(y = "Collection Temp. (°C)",
       x = "") + 
  theme_matt_facets(base_size = 14)

sal_comp_sal = sal_comps %>%  
  select(salinity, season, region, collection_temp, collection_salinity) %>% 
  distinct() %>% 
  ggplot(aes(x = salinity, y = collection_salinity, colour = season, group = season)) + 
  facet_wrap(region~.) + 
  geom_line(linewidth = 1.5) + 
  geom_point(size = 4) + 
  scale_colour_manual(values = season_cols) + 
  labs(y = "Collection Salinity (psu)",
       x = "Salinity") + 
  theme_matt_facets(base_size = 14)

ggarrange(sal_comp_temps, sal_comp_sal, nrow = 2, common.legend = T, legend = "right")
```

The latitudinal gradient covers a wide range of seasonality. Shown below is the temperature range. While based on collection temperatures, and therefore an underestimate of the total seasonal range of temperatures, these patterns are representative of the expected latitudinal gradient in seasonality. 

```{r lat-temp-range-plot, fig.width=8, fig.height=5}
site_temps %>% 
  group_by(site, lat) %>%  
  summarise(temp_range = max(collection_temp) - min(collection_temp)) %>%  
  ggplot(aes(x = lat, y = temp_range)) + 
  geom_point(aes(colour = site),
             size = 3) + 
  scale_color_manual(values = site_cols) + 
  labs(x = "Latitude", 
       y = "Collection Temp. Range (°C)") + 
  theme_matt() + 
  theme(legend.position = "right")
```


## Phenotypic Measurements 

### Critical Thermal Limits

A total of `r dim(all_data)[1]` individuals were examined. Critical thermal limits and body size measurements were made before individuals were preserved in ethanol. We excluded data for `r dim(excluded)[1]` individuals, detailed below. These individuals had either very low CTmax or were, upon re-examination of photographs, identified as juveniles instead of mature females. With these individuals excluded, **the full data set contains `r dim(full_data)[1]` phenotyped individuals**.   

```{r excluded-inds}
excluded %>% 
  select(region, site, season, collection_temp, collection_salinity, replicate, tube, ctmax, size) %>% 
  knitr::kable(align = "c")
```

Critical thermal maxima (CTmax) was measured using a custom setup. The method uses a standard dynamic ramping assay to determine the maximum temperature individuals could sustain normal functioning. This differs from lethal temperatures, and indeed, all individuals observed recovered following the assay.    

Individuals were rested for one hour after collection before the assay. During the assay, copepods were held in artificial seawater, composed of bottled spring water and Instant Ocean salt mix adjusted to match collection salinities. During the assay, several 'control' individuals were maintained in this solution at ambient temperatures without the temperature ramp to ensure that there was no background mortality. When sorting individuals from the plankton tow contents, they were held in a 50:50 mix of 60 um filtered water from the collection site and artificial seawater as an additional acclimation step. 

Sample sizes varied slightly across experiments, but generally ~20 individuals were measured per season at each site. The major exceptions were the early samples from the Florida sites and the late sample from Sawyer Park (Maine). Only two sets of samples (peak and late) were collected from Fort Hamer and Manatee River. The late season collection from Sawyer Park occurred after *Acartia tonsa* abundance decreased, and samples were dominated by *Acartia hudsonica*. Thermal limits were measured for all *A. tonsa* individuals recovered. No samples were collected from Key Largo for this project, as *Acartia tonsa* wasn't present in the water during the peak season, likely due to a recent extreme heat event. Thermal limits from a separate project that examined thermal limits at this site (collected February 2023; Sasaki et al. 2023) are included as a point of comparison. Body size values were not measured for these Key Largo copepods, nor were copepods individually preserved after the experiments.   

```{r}
full_data %>%  
  group_by(site, season) %>% 
  count() %>%  
  pivot_wider(id_cols = "site", 
              names_from = season, 
              values_from = n, 
              values_fill = 0) %>% 
  knitr::kable()
```

Shown below are the measured CTmax values. Individual measurements are shown in small points for each collection. The large points indicate the mean values for each collection.     

```{r seasonal-ct-max}
mean_ctmax = full_data %>% 
  group_by(site, season, doy, collection_temp) %>% 
  summarize(mean_ctmax = mean(ctmax),
            median_ctmax = median(ctmax))

ctmax_plot = ggplot(full_data, aes(x = season, y = ctmax, colour = site)) + 
  geom_line(data = mean_ctmax, 
            aes(y = mean_ctmax, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 1) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.3) + 
  geom_point(data = mean_ctmax, 
             aes(y = mean_ctmax),
             position = position_dodge(width = 0.4),
             size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "right", 
        legend.title.align = 0.125)

ctmax_plot
```

```{r doy-ct-max, include = F}
# The same data is shown below, plotted against day of the year instead of season. This accounts for the variable timing of collections across regions (e.g. - the compressed collections from the Northern sites to accomodate the earlier onset of cold temperatures). 

ggplot(filter(full_data, site != "Key Largo"), aes(x = doy, y = ctmax, colour = site)) + 
  geom_line(data = filter(mean_ctmax, site != "Key Largo"), 
            aes(y = median_ctmax, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 1) + 
  geom_point(position = 
               position_jitterdodge(
                 jitter.width = 0.1, 
                 jitter.height = 0,
                 dodge.width = 0.4),
             alpha = 0.3) + 
  geom_point(data = filter(mean_ctmax, site != "Key Largo"), 
             aes(y = median_ctmax),
             position = position_dodge(width = 0.4),
             size = 4) + 
  scale_colour_manual(values = site_cols) + 
  xlim(180,350) + 
  labs(y = "CTmax (°C)",
       x = "Day of the Year") +
  theme_matt() + 
  theme(legend.position = "right", 
        legend.title.align = 0.125)
```

```{r ctmax-ind-pops-doy, fig.width=8, fig.height=6, include = F}
# CTmax data for each individual site is shown below, plotted against day of the year. 

ggplot(filter(full_data, site != "Key Largo"), 
       aes(x = doy, y = ctmax, colour = site)) + 
  facet_wrap(.~site, scales = "free") + 
  geom_line(data = filter(mean_ctmax, site != "Key Largo"), 
            aes(y = mean_ctmax, group = site),
            linewidth = 3, alpha = 0.5) + 
  geom_point(alpha = 0.8) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Day of Year") +
  theme_matt() + 
  theme(legend.position = "none", 
        legend.title.align = 0.125,
        axis.text.x = element_text(angle = 320, hjust = 0, vjust = 0.5))
```

```{r ctmax-ind-pops-season, fig.width=8, fig.height=6, include = F}
ggplot(full_data, aes(x = season, y = ctmax, colour = site)) + 
  facet_wrap(.~site, scales = "free_y") + 
  geom_line(data = mean_ctmax, 
            aes(y = median_ctmax, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 3, alpha = 0.5) + 
  geom_line(data = mean_ctmax, 
            aes(y = collection_temp, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 2,
            colour = "grey") + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.8) + 
  # geom_point(data = mean_ctmax, 
  #            aes(y = median_ctmax),
  #            position = position_dodge(width = 0.4),
  #            size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "none", 
        legend.title.align = 0.125)
```

### Warming tolerance

Warming tolerance (the difference between thermal limits and environmental temperatures) is a commonly used metric of climate vulnerability. We calculated this as the difference between measured CTmax values and the collection temperature. Smaller warming tolerance values indicate that populations were nearer to their upper thermal limits, suggesting greater vulnerability. 

```{r seasonal-warming-tol}
mean_wt = full_data %>% 
  group_by(site, season) %>% 
  summarize(mean_wt = mean(warming_tol),
            median_wt = median(warming_tol))

ggplot(full_data, aes(x = season, y = warming_tol, colour = site)) + 
  geom_line(data = mean_wt, 
            aes(y = mean_wt, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 1) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.3) + 
  geom_point(data = mean_wt, 
             aes(y = mean_wt),
             position = position_dodge(width = 0.4),
             size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Warming Tolerance (°C)",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "right", 
        legend.title.align = 0.125)
```

### Body Size

Following the CTmax assay, individuals were photographed for body size measurements. Prosome lengths were measured from these photographs using a scale micrometer and the software ImageJ. These measurements are shown below. As before, large points indicate the mean body size. While less cohesive than CTmax, a general trend of increasing body size with latitude and time of year can be seen. 

```{r seasonal-body-size}
mean_size = full_data %>% 
  group_by(site, season, doy, collection_temp) %>% 
  summarize(mean_size = mean(size),
            median_size = median(size))

ggplot(full_data, aes(x = season, y = size, colour = site)) + 
  geom_line(data = mean_size, 
            aes(y = mean_size, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 1) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.3) + 
  geom_point(data = mean_size, 
             aes(y = mean_size),
             position = position_dodge(width = 0.4),
             size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "right", 
        legend.title.align = 0.125)
```

```{r doy-body-size, include = F}
ggplot(drop_na(full_data, size), aes(x = doy, y = size, colour = site)) + 
  geom_line(data = drop_na(mean_size, mean_size), 
            aes(y = median_size, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 1) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.3) + 
  geom_point(data = drop_na(mean_size, mean_size), 
             aes(y = median_size),
             position = position_dodge(width = 0.4),
             size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Day of the Year") +
  theme_matt() + 
  theme(legend.position = "right", 
        legend.title.align = 0.125)
```


```{r size-ind-pops-doy, fig.width=8, fig.height=6, include = F}
# Shown below is the body size data for each individual site. 

ggplot(drop_na(full_data, size), aes(x = doy, y = size, colour = site)) + 
  facet_wrap(.~site) + 
  geom_line(data = drop_na(mean_size, mean_size), 
            aes(y = mean_size, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 3, alpha = 0.5) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.8) + 
  # geom_point(data = mean_ctmax, 
  #            aes(y = median_ctmax),
  #            position = position_dodge(width = 0.4),
  #            size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Day of Year") +
  theme_matt() + 
  theme(legend.position = "none", 
        legend.title.align = 0.125)
```

```{r size-ind-pops-season, fig.width=8, fig.height=6, include = F}
ggplot(full_data, aes(x = season, y = size, colour = site)) + 
  facet_wrap(.~site) + 
  geom_line(data = mean_size, 
            aes(y = median_size, group = site),
            position = position_dodge(width = 0.4),
            linewidth = 3, alpha = 0.5) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.8) + 
  # geom_point(data = mean_ctmax, 
  #            aes(y = median_ctmax),
  #            position = position_dodge(width = 0.4),
  #            size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "none", 
        legend.title.align = 0.125)
```

### Salinity Pair Comparisons 

The three pairs of cross-salinity comparisons do show evidence for fine-scale trait divergence, although there was no consistent pattern in the direction or magnitude of differences. CTmax was similar across sites in the Southern and Central pairs. In the Northern pair, CTmax tended to be slightly lower for individuals from the low salinity site. Size was more variable between the paired sites. In the South, individuals from the low salinity site were consistently smaller than individuals from the high salinity site, despite experiencing similar temperatures. By contrast, in the Central pair individuals from the low salinity site tended to be slightly larger than those from the high salinity site, although the magnitude of this difference this varied seasonally. In the Northern pair, individuals from the low salinity site were slightly larger than individuals from the high salinity site during the early and peak season collections, but there was an abrupt reversal of this pattern during the late collection, when individuals from the high salinity site were much larger than individuals from the low salinity site.

We will return to potential mechanisms behind divergence between these geographically proximate but environmentally separated sites later in this report. 

```{r sal-pair-traits}
sal_means = sal_comps %>% 
  group_by(region, site, salinity, season) %>% 
  summarize(mean_ctmax = mean(ctmax),
            mean_size = median(size))

sal_comp_ctmax_plot = sal_comps %>% 
  ggplot(aes(x = salinity, y = ctmax, colour = season, group = season)) + 
  facet_wrap(region~.) + 
    geom_line(data = sal_means, 
            aes(y = mean_ctmax, group = season),
            position = position_dodge(width = 0.4),
            linewidth = 1) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.3) + 
  geom_point(data = sal_means, 
             aes(y = mean_ctmax),
             position = position_dodge(width = 0.4),
             size = 4) + 
  scale_colour_manual(values = season_cols) + 
  labs(y = "CTmax (°C)",
       x = "") + 
  theme_matt_facets(base_size = 14)

sal_comp_size_plot = sal_comps %>% 
  ggplot(aes(x = salinity, y = size, colour = season, group = season)) + 
  facet_wrap(region~.) + 
    geom_line(data = sal_means, 
            aes(y = mean_size, group = season),
            position = position_dodge(width = 0.4),
            linewidth = 1) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.4),
             alpha = 0.3) + 
  geom_point(data = sal_means, 
             aes(y = mean_size),
             position = position_dodge(width = 0.4),
             size = 4) + 
  scale_colour_manual(values = season_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "") + 
  theme_matt_facets(base_size = 14)

ggarrange(sal_comp_ctmax_plot, sal_comp_size_plot, nrow = 2, common.legend = T, legend = "right")

```

## Trait Correlations

Regardless of the underlying mechanism (genetic differentiation or phenotypic plasticity), we expect that collections from warmer waters should yield copepods with higher thermal limits and smaller body sizes. Our observations largely fit this expectation, with strong increases in CTmax at higher temperatures, and a general decrease in prosome lengths as temperature increased. There was a larger correlation between CTmax and temperature than size and temperature. There are myriad reasons this might be observed, but one potential is the potential for acclimation of thermal limits. Body size is fixed in the adult stage, while CTmax can continue to vary as water temperature changes. This is indirect evidence for the importance of rapid acclimation in the observed patterns. 

```{r temp-cors, fig.width=15, fig.height=6}
ctmax_temp_plot = ggplot(full_data, aes(x = collection_temp, y = ctmax)) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2, 
              colour = "grey") + 
  geom_point(aes(colour = site), 
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "none")

size_temp_plot = ggplot(full_data, aes(x = collection_temp, y = size)) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2, 
              colour = "grey") + 
  geom_point(aes(colour = site), 
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

wt_temp_plot = ggplot(full_data, aes(x = collection_temp, y = warming_tol)) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2, 
              colour = "grey") + 
  geom_point(aes(colour = site), 
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Warming Tolerance (°C)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

ggarrange(ctmax_temp_plot, wt_temp_plot, size_temp_plot, common.legend = T, legend = "bottom", nrow = 1)
```

```{r pop-temp-cors, fig.width=8, fig.height=6, include = F}
# Shown below are these temperature correlations for each individual population. Variation in the temperature sensitivity of these traits appears to vary across populations, with reduced slopes in both the lowest and the highest latitude populations. Again, we emphasize that this observed trait variation may stem from either (or both) plastic and genetic effects. However, these observations provide estimates for realistic patterns in temperature sensitivity for these populations. 

ggplot(full_data, aes(x = collection_temp, y = ctmax)) + 
  facet_wrap(.~site) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2, 
              colour = "grey") + 
  geom_point(aes(colour = site), 
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "none")

ggplot(full_data, aes(x = collection_temp, y = size)) + 
  facet_wrap(.~site) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2, 
              colour = "grey") + 
  geom_point(aes(colour = site), 
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "none")

```

The relationship between prosome length and CTmax is also of interest. Given that larger body sizes are associated with cold adaptation/acclimation, it is unsurprising that there is a general trend of decreasing thermal limits with increasing size. Individual regression lines for each site are also included - the dark grey lines in the background represent the 'universal' regression for that site, with individual colored regression lines for each collection. At most sites, evidence for this negative relationship between CTmax and body size is observed across collections rather than within collections, suggesting this relationship is primaly a reflection of the response to seasonal temperature variation. 

```{r ctmax-vs-size, fig.width=6, fig.height=10}
universal_size = full_data %>% 
  ggplot(aes(x = size, y = ctmax)) + 
  # geom_smooth(data = filter(full_data, ctmax > 31), 
  #             aes(x = size, y = ctmax),
  #             method = "lm", 
  #             colour = "grey60", 
  #             se = F,
  #             linewidth = 2) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2,
              colour = "grey70") + 
  geom_point(aes(colour = site),
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "") +
  theme_matt(base_size = 14) + 
  theme(legend.position = "right",
        axis.title.x = element_blank())

pop_size = full_data %>% 
  ggplot(aes(x = size, y = ctmax, colour = site, group = season)) + 
  facet_wrap(site~.) + 
  # geom_smooth(data = filter(full_data, ctmax > 31), 
  #             aes(x = size, y = ctmax),
  #             method = "lm", 
  #             colour = "grey60", 
  #             se = F,
  #             linewidth = 2) + 
  geom_smooth(data = full_data, 
              aes(x = size, y = ctmax, group = site), 
              colour = "grey20", method = "lm", se = F) + 
  geom_point(size = 1.3, alpha = 0.3) + 
  geom_smooth(method = "lm", se = F,
              linewidth = 1) + 
  scale_colour_manual(values = site_cols) + 
  scale_x_continuous(breaks = c(0.6, 0.8, 1)) + 
  labs(y = "CTmax (°C)",
       x = "Prosome Length (mm)") +
  theme_matt(base_size = 14) + 
  theme(legend.position = "right")

ggarrange(universal_size, pop_size, common.legend = T, legend = "none", nrow = 2)
```

If there is a true mechanistic relationship between body size and thermal limits, we would expect to see this relationship emerge **within** populations, or even individual collections. Shown below is the relationship between CTmax and size residuals, acquired from regressions of these traits against collection temperature. This substantially reduces the strength of the apparent relationship, but there is still a slightly negative overall relationship, spanning both across-population, within-population, and even within-collection scales (see the Sawyer Park collections, for example).     

```{r ctmax-vs-size-resids, fig.width=6, fig.height=10}
filtered_data = full_data %>% 
  ungroup() %>% 
  drop_na(size, ctmax) %>% 
  mutate(temp_cent = scale(collection_temp, scale = F)[,1],
         size_cent = scale(size, scale = F)[,1],
         sal_cent = scale(collection_salinity, scale = F)[,1],
         sal_type = if_else(collection_salinity > 15, "High", "Low"))

ctmax_temp.model = lm(ctmax ~ collection_temp + site, data = filtered_data)
ctmax_resids = residuals(ctmax_temp.model)

size_temp.model = lm(size ~ collection_temp + site, data = filtered_data)
size_resids = residuals(size_temp.model)

universal_resids = filtered_data %>% 
  mutate(ctmax_resids = ctmax_resids,
         size_resids = size_resids) 

all_resids = ggplot(universal_resids, aes(x = size_resids, y = ctmax_resids)) + 
  # geom_smooth(data = filter(full_data, ctmax > 31), 
  #             aes(x = size, y = ctmax),
  #             method = "lm", 
  #             colour = "grey60", 
  #             se = F,
  #             linewidth = 2) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2,
              colour = "grey70") + 
  geom_point(aes(colour = site),
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax Residuals",
       x = "") +
  theme_matt(base_size = 14) + 
  theme(legend.position = "right",
        axis.title.x = element_blank())

pop_resids = ggplot(universal_resids, aes(x = size_resids, y = ctmax_resids, colour = site, group = season)) + 
  facet_wrap(site~.) + 
  # geom_smooth(data = filter(full_data, ctmax > 31), 
  #             aes(x = size, y = ctmax),
  #             method = "lm", 
  #             colour = "grey60", 
  #             se = F,
  #             linewidth = 2) + 
  geom_smooth(aes(x = size_resids, y = ctmax_resids, group = site), 
              colour = "grey20", method = "lm", se = F) + 
  geom_point(size = 1.3, alpha = 0.3) + 
  geom_smooth(method = "lm", se = F,
              linewidth = 1) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax Residuals",
       x = "Prosome Length Residuals") +
  theme_matt(base_size = 14) + 
  theme(legend.position = "right")

ggarrange(all_resids, pop_resids, common.legend = T, legend = "none", nrow = 2)
```

```{r ctmax-model, include = F}
ctmax.model = lme4::lmer(
  data = filtered_data, 
  ctmax ~ temp_cent + size + sal_type + (1 + temp_cent|site))

lmerTest::ranova(ctmax.model)
car::Anova(ctmax.model)

```

To more formally test the relationships between CTmax, collection temperature, and size, we used a linear mixed effects model, structured as `ctmax ~ collection temp. + size + salinity + (1 + collection temp.|site)`. This examines the effects of temperature and size on CTmax, along with differences between the salinity groupings. Collection temperature and size were both centered and salinity transformed into a categorical variable ("low" salinity as anything below 15 psu, "high" salinity as anything above 20 psu) before model fitting. The model also includes random intercepts for each site and random slopes for collection temperature (i.e. - variation in the acclimation capacity of CTmax). 

All three factors (collection temperature, body size, and salinity type) had a significant effect on CTmax. The overall effect of temperature suggests an increase in CTmax of `r round(unname(fixef(ctmax.model)["temp_cent"]), digits = 2)`°C per °C increase in collection temperature (i.e. - an ARR value of `r round(unname(fixef(ctmax.model)["temp_cent"]), digits = 2)`), while increasing body sizes decrease CTmax by `r round(unname(fixef(ctmax.model)["size"]), digits = 2)`°C per mm (or a decrease of ~`r round(unname(fixef(ctmax.model)["size"]), digits = 2)/10`°C per tenth of a mm, which is more biologically realistic for *A. tonsa*). The model suggests copepods from low salinity sites had lower thermal limits than copepods from high salinity sites by ~`r abs(round(unname(fixef(ctmax.model)["sal_typeLow"]), digits = 1))`°C.

```{r}

effects_summary = data.frame(
  "Temperature" = unname(fixef(ctmax.model)["temp_cent"]),
  "Size" = unname(fixef(ctmax.model)["size"]),
  "Salinity" = unname(fixef(ctmax.model)["sal_typeLow"])) %>%  
  select("Temperature (per degree)" = Temperature, 
         "Size (per mm)" = Size, 
         "Salinity (Low vs. High)" = Salinity)

knitr::kable(effects_summary)
```

By extracting the conditional mode for the random effects, we can also examine how thermal limits vary across sites beyond the influence of collection temperatures and body sizes. Shown below are these "population" values (essentially, all else being equal, how would thermal limit vary across populations). We can see that, similar to what's been observed in common garden experiments with *A. tonsa* previously, there is a general latitudinal trend in thermal limits (decreasing with increasing latitude). Divergence is particularly strong in the Northern populations.    

```{r pop-effs-plot, fig.width=10, fig.height=5}
pop_effs = REsim(ctmax.model) %>% 
  dplyr::select("site" = groupID, term, mean, sd) %>% 
  filter(term == "(Intercept)") %>% 
  inner_join(temp_summaries, by = c("site")) %>% 
  mutate(site = fct_reorder(site, lat))

season_effs =season_effs =season_effs = REsim(ctmax.model) %>% 
  dplyr::select("site" = groupID, term, mean, sd) %>% 
  filter(site %in% c("early", "peak", "late")) %>% 
  select("season" = site, mean, sd)

#plotREsim(REsim(ctmax.model))  # plot the interval estimates

pop_effs_plot = ggplot(pop_effs, aes(x = lat, y = mean, colour = site)) + 
  geom_hline(yintercept = 0, colour = "grey") +
  geom_errorbar(aes(ymin = mean - 1.96 * sd, ymax = mean + 1.96 * sd),
                width = 0.5, linewidth = 1) + 
  geom_point(size = 3) + 
  scale_colour_manual(values = site_cols) + 
  labs(x = "Latitude", 
       y = "Population Effect") + 
  theme_matt() + 
  theme(legend.position = "right")

pop_clim_plot = ggplot(pop_effs, aes(x = season_mean, y = mean)) + 
  #geom_abline(slope = 1, intercept = -19) + 
  geom_smooth(method = "lm", colour = "black") +
  geom_point(aes(colour = site), size = 4) + 
  scale_colour_manual(values = site_cols) + 
  labs(x = "Mean Season Temp.", 
       y = "Population Effect") + 
  theme_matt() + 
  theme(legend.position = "right")

ggarrange(pop_effs_plot, pop_clim_plot, common.legend = T, legend = "bottom")
```

Finally, shown below are the estimated random slopes for each site. These represent the effects of collection temperature on CTmax for each site. Interestingly, these estimates diverge from the results of previous common garden experiments, which showed the strongest plasticity in high latitude sites. Here, acclimation appears to peak in mid-latitudes, and decrease at both high and low latitude sites. This suggests factors may constrain acclimation capacity in natural populations (e.g. food availability, environmental variation, pathogens, etc.).    

```{r site-arr-plot, fig.width=6, fig.height=4}
coefficients(ctmax.model)$site %>%
  janitor::clean_names() %>%
  rownames_to_column(var = "site") %>%
  inner_join(site_data) %>% 
  mutate(site = fct_reorder(site, lat)) %>% 
  ggplot(aes(x = temp_cent, y = site)) +
  geom_point(aes(colour = site),
             size = 5) +
  scale_colour_manual(values = site_cols) +
  labs(x = "ARR") + 
  theme_matt() +
  theme(legend.position = "none",
        axis.title.y = element_blank())
```

## Trait Variability

Shown below is the trait variation (ranges) for each site. Ranges are calculated for each collection separately. 

```{r trait-range-plot, fig.width=12, fig.height=4.5}
trait_ranges = full_data %>% 
  group_by(site, season, collection_temp, collection_salinity, doy, lat) %>% 
  summarise(mean_ctmax = mean(ctmax),
            ctmax_range = max(ctmax) - min(ctmax),
            ctmax_var = var(ctmax),
            mean_size = mean(size),
            size_range = max(size) - min(size),
            size_var = var(size)) %>% 
  mutate(prop_ctmax_range = ctmax_range / mean_ctmax,
         prop_size_range = size_range / mean_size)

ctmax_range_temp = ggplot(trait_ranges, aes(x = collection_temp, y = ctmax_range)) + 
  geom_smooth(method = "lm", colour = "black") + 
  geom_point(aes(colour = site), 
             size = 3) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax Range (°C)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

ctmax_var_temp = ggplot(trait_ranges, aes(x = collection_temp, y = ctmax_var, colour = site)) + 
  geom_point(size = 3) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax Range (°C)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

size_range_temp = ggplot(trait_ranges, aes(x = collection_temp, y = size_range, colour = site)) + 
  geom_point(size = 3) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Size Range (mm)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

size_var_temp = ggplot(trait_ranges, aes(x = collection_temp, y = size_var, colour = site)) + 
  geom_point(size = 3) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Size Range (mm)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

ggarrange(ctmax_range_temp, size_range_temp, common.legend = T, legend = "right")
```

Changes in trait variance may be indicative of phenotypic selection. If selection (as opposed to acclimation) are driving seasonal changes, we may expect to see a reduction in variance in the peak samples relative to the early season samples. Note that early season collection temperatures this year were higher than expected, driven by fairly strong heatwaves across the North Atlantic. As shown in the temperature profiles for each site, the 'early' samples were collected just after high temperatures were reached, while 'peak' samples were collected after sites had experienced high temperatures for several weeks (generations). As warming tolerances were fairly high throughout this period, we will assume that selection was weak before the early samples. If the early onset of high temperatures filtered out vulnerable genotypes prior to our sampling, the results will be a conservative estimate of the effects of selection on trait variance.     

Shown below is the seasonal progression of variance in CTmax for each site. For many sites, variance decreased between the early and peak samples, and then increased again in the late sample. For some sites (e.g. Esker Point), this increase in the late sample was substantial. 

```{r season-var}
ggplot(trait_ranges, aes(x = season, y = ctmax_var, colour = site)) + 
  geom_line(aes(group = site), 
            linewidth = 1.5) + 
  geom_point(size = 3) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax Variance",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "right", 
        legend.title.align = 0.125)
```

Shown below is the seasonal progression of variance in body size. A similar pattern of decreasing variance in peak samples relative to early and late samples is again seen for many sites. The obvious exception is the Esker Point sample, which saw the opposite trend. 

```{r}
ggplot(trait_ranges, aes(x = season, y = size_var, colour = site)) + 
  geom_line(aes(group = site), 
            linewidth = 1.5) + 
  geom_point(size = 3) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax Variance",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "right", 
        legend.title.align = 0.125)
```

## Low-Coverage Whole Genome Sequencing   

While plasticity appears to play a large role in the response of *Acartia tonsa* populations to seasonality, a growing number of studies have also uncovered patterns in allele frequency indicating of rapid adaptive responses to seasonality. Our data set of seasonal samples across a large latitudinal gradient enables a comparison of these fine scale temporal patterns with the larger latitudinal patterns in allele frequency to determine whether the same alleles driving rapid seasonal adaptation are in play over larger spatial (and longer temporal) scales.

After phenotyping, each individual was preserved in 95% ethanol. Individual DNA libraries were prepared using Twist Bio 96-plex prep kits, then sequenced on two lanes of an Illumina NovaSeq X Plus 25B by Novogene. Reads were de-multiplexed first to the plate level, and then to the sample level. The number of reads per sample varied, as shown below. 

```{r overall-read-counts, fig.width=6, fig.height=6}
read_data  %>% 
  mutate(site = fct_relevel(site, "Manatee River", "Ft. Hamer", 
                            "Tyler Cove", "Ganey's Wharf", "Esker Point", 
                            "Sawyer Park", "St. Thomas de Kent Wharf", "Ritchie Wharf"),
         season = fct_relevel(season, "early", "peak", "late"),
         reads_millions = templates / 1000000) %>% 
  ggplot(aes(reads_millions))+ 
  geom_histogram(binwidth = 2) + 
  theme_matt()
```

The sample read counts are shown below separated out by site and season. 

```{r site-season-read-counts, fig.width=9, fig.height=14}
read_data %>% 
  mutate(site = fct_relevel(site, "Manatee River", "Ft. Hamer", 
                            "Tyler Cove", "Ganey's Wharf", "Esker Point", 
                            "Sawyer Park", "St. Thomas de Kent Wharf", "Ritchie Wharf"),
         season = fct_relevel(season, "early", "peak", "late"),
         reads_millions = templates / 1000000) %>% 
  ggplot(aes(reads_millions, fill = site)) + 
  facet_grid(site~season) + 
  geom_histogram(binwidth = 10) + 
  scale_fill_manual(values = site_cols) +
  scale_y_continuous(breaks = c(0, 5, 10)) + 
  scale_x_continuous(breaks = c(0, 40)) + 
    theme_matt_facets() + 
  theme(legend.position = "none")
```

While the number of reads recovered for each individual varies, there is no clear relationship between read counts and size or CTmax. 

```{r}
pheno_read_comps = full_data %>%  
  dplyr::select(site, season, replicate, tube, size, ctmax) %>% 
  inner_join(read_data)

reads_size_plot = ggplot(pheno_read_comps, aes(x = size, y = templates)) + 
  geom_smooth() + 
  geom_point() + 
  theme_matt()

reads_ctmax_plot = ggplot(pheno_read_comps, aes(x = ctmax, y = templates)) + 
  geom_smooth() + 
  geom_point() + 
  theme_matt()

ggarrange(reads_size_plot, reads_ctmax_plot, nrow = 1)

read.model = lm(data = pheno_read_comps, 
                templates ~ size + ctmax)

knitr::kable(car::Anova(read.model))

```

*Acartia tonsa* is well known to comprise multiple, deeply diverged clades. To determine the clade composition of our sample set, we skimmed COI sequences from the low-coverage data. 

```{r fig.width=15, fig.height=9}
# clade_matches %>% 
#   group_by(sample) %>% 
#   count(Clade) %>% 
#   mutate(population = str_split_fixed(sample, pattern = "_", n = 4)[,1],
#          season = str_split_fixed(sample, pattern = "_", n = 4)[,2],
#          individual = str_split_fixed(sample, pattern = "_", n = 3)[,3])  %>% 
#   ungroup() %>% 
#   mutate(population = fct_relevel(population, "MR", "FH", "MD", "GW", "CT", "ME", "TK", "RW"), 
#          season = fct_relevel(season, "early", "peak", "late"), 
#          sample = fct_reorder2(sample, .y = population, .x = season, .desc = F))
#   #write.csv(file = "Output/Output_data/COI_clades_summary.csv", row.names = F)

ggplot(clade_summary, aes(x = individual, y = n, fill = Clade)) + 
  facet_grid(population~season) + 
  geom_bar(stat = "identity", position = "fill") + 
  scale_y_continuous(breaks = c(0, 0.5, 1)) + 
  scale_fill_manual(values = c("A_hudsonica" = "#1B9E99",
                               "A_lilljeborgi" = "#D95F02",
                               "F" = "#7570B3",
                               "IV" = "#E7298A",
                               "out_group" = "#666666",
                               "S" = "#66A61E",
                               "SB" = "#E6AB02",
                               "X" = "#A6761D"),
                    na.value = "black") + 
  labs(x = "Individual", 
       y = "Proportion") + 
  theme_bw(base_size = 18) + 
  theme(panel.grid = element_blank(), 
        axis.text.x = element_text(angle = 290, hjust = 0, vjust = 0.5))
```

These COI clades were then merged with the CTmax and body size data set to examine clade-specific patterns of divergence. By plotting CTmax and size against the collection temperature we can see that these clades appear to be physiologically differentiated. Clades F and S tend to have higher CTmax, even when collected at the same temperatures as Clades X and IV. Size is more variable, with varying degrees of temperature sensitivity across the clades. Overall, however, Clade F tends to be smaller than Clades S and X. Clade IV tended to be the largest. 

```{r fig.width=11, fig.height=6}
clade_ctmax_plot = ggplot(clade_ctmax, aes(x = collection_temp, y = ctmax, colour = Clade)) + 
  geom_point(size = 2, alpha = 0.5) + 
  geom_smooth(method = "lm", linewidth = 2) + 
  labs(x = "Collection Temp. (°C)", 
       y = "CTmax (°C)") + 
  scale_colour_manual(values = c("F" = "#7570B3",
                               "IV" = "#E7298A",
                               "S" = "#66A61E",
                               "X" = "#A6761D")) + 
  theme_matt()

clade_size_plot = ggplot(clade_ctmax, aes(x = collection_temp, y = size, colour = Clade)) + 
  geom_point(size = 2, alpha = 0.5) + 
  geom_smooth(method = "lm", linewidth = 2) + 
    labs(x = "Collection Temp. (°C)", 
       y = "Prosome Length (mm)") + 
  scale_colour_manual(values = c("F" = "#7570B3",
                               "IV" = "#E7298A",
                               "S" = "#66A61E",
                               "X" = "#A6761D")) + 
  theme_matt()

ggarrange(clade_ctmax_plot, clade_size_plot, nrow = 1, common.legend = T, legend = "bottom")

# ggplot(clade_ctmax, aes(x = size, y = ctmax, colour = Clade)) + 
#   geom_point(size = 2) + 
#   #geom_smooth(method = "lm", linewidth = 2) + 
#     labs(y = "CTmax (°C)", 
#        x = "Prosome Length (mm)") + 
#   scale_colour_manual(values = c("A_hudsonica" = "#1B9E99",
#                                "A_lilljeborgi" = "#D95F02",
#                                "F" = "#7570B3",
#                                "IV" = "#E7298A",
#                                "out_group" = "#666666",
#                                "S" = "#66A61E",
#                                "SB" = "#E6AB02",
#                                "X" = "#A6761D"),
#                     na.value = "black") + 
#   theme_matt_facets()

```

These differences in physiology correspond to the distributional patterns for each clade - shown below are logistic regressions of clade presence against latitude or collection temperature. 

```{r fig.width=12, fig.height=6}
clade_occurence = clade_ctmax %>% 
  filter(Clade != "A_hudsonica") %>% 
  select(site, season, Clade, collection_temp, lat) %>% 
  group_by(site, season, collection_temp, lat) %>% 
  count(Clade) %>% 
  ungroup() %>% 
  complete(nesting(site, season, collection_temp, lat), Clade, fill = list(n = 0)) %>%  
  mutate(n = if_else(n == 0, 0, 1), 
         num_season = case_when(
           season == "early" ~ 1,
           season == "peak" ~ 2,
           season == "late" ~ 3))

lat_clade_plot = ggplot(clade_occurence, aes(x = lat, y = n, colour = Clade)) + 
  facet_wrap(Clade~.) + 
  geom_point(size = 3, 
             position = position_jitter(width = 0, height = 0.02)) + 
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = T, 
    linewidth = 3) + 
  labs(x = "Latitude", 
       y = "Presence") + 
  scale_y_continuous(breaks = c(0,1)) + 
  scale_colour_manual(values = c("A_hudsonica" = "#1B9E99",
                               "A_lilljeborgi" = "#D95F02",
                               "F" = "#7570B3",
                               "IV" = "#E7298A",
                               "out_group" = "#666666",
                               "S" = "#66A61E",
                               "SB" = "#E6AB02",
                               "X" = "#A6761D"),
                    na.value = "black") + 
  theme_matt_facets() + 
  theme(legend.position = "none")

temp_clade_plot = ggplot(clade_occurence, aes(x = collection_temp, y = n, colour = Clade)) + 
  facet_wrap(Clade~.) + 
  geom_point(size = 3, 
             position = position_jitter(width = 0, height = 0.02)) + 
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = T, 
    linewidth = 3) + 
  labs(x = "Collection Temp. (°C)", 
       y = "Presence") + 
  scale_y_continuous(breaks = c(0,1)) + 
  scale_colour_manual(values = c("A_hudsonica" = "#1B9E99",
                               "A_lilljeborgi" = "#D95F02",
                               "F" = "#7570B3",
                               "IV" = "#E7298A",
                               "out_group" = "#666666",
                               "S" = "#66A61E",
                               "SB" = "#E6AB02",
                               "X" = "#A6761D"),
                    na.value = "black") + 
  theme_matt_facets() + 
  theme(legend.position = "none")

ggarrange(lat_clade_plot, temp_clade_plot)

```

These relationships are refined a bit when switching to quasi-binomial regressions for the clades' proportional abundance (what proportion of individuals in each collection were assigned to a clade). 

```{r fig.width=12, fig.height=6}
clade_prop = clade_ctmax %>% 
  filter(Clade != "A_hudsonica") %>% 
  select(site, season, Clade, collection_temp, lat) %>% 
  group_by(site, season, collection_temp, lat) %>% 
  count(Clade) %>% 
  ungroup(collection_temp, lat) %>% 
  mutate("prop" = n / sum(n)) %>% 
  ungroup() %>% 
  complete(nesting(site, season, collection_temp, lat), Clade, fill = list(prop = 0, n = 0)) %>%  
  mutate(n = if_else(n == 0, 0, 1), 
         num_season = case_when(
           season == "early" ~ 1,
           season == "peak" ~ 2,
           season == "late" ~ 3))


lat_clade_prop_plot = ggplot(clade_prop, aes(x = lat, y = prop, colour = Clade)) + 
  facet_wrap(Clade~.) + 
  geom_point(size = 3) + 
  geom_smooth(method = "glm", 
    method.args = list(family = "quasibinomial"), 
    se = T, 
    linewidth = 3) + 
  labs(x = "Latitude", 
       y = "Sample Proportion") + 
  scale_y_continuous(breaks = c(0,1)) + 
  scale_colour_manual(values = c("A_hudsonica" = "#1B9E99",
                               "A_lilljeborgi" = "#D95F02",
                               "F" = "#7570B3",
                               "IV" = "#E7298A",
                               "out_group" = "#666666",
                               "S" = "#66A61E",
                               "SB" = "#E6AB02",
                               "X" = "#A6761D"),
                    na.value = "black") + 
  theme_matt_facets() + 
  theme(legend.position = "none")

temp_clade_prop_plot = ggplot(clade_prop, aes(x = collection_temp, y = prop, colour = Clade)) + 
  facet_wrap(Clade~.) + 
  geom_point(size = 3) + 
  geom_smooth(method = "glm", 
    method.args = list(family = "quasibinomial"), 
    se = T, 
    linewidth = 3) + 
  labs(x = "Collection Temp. (°C)", 
       y = "Sample Proportion") + 
  scale_y_continuous(breaks = c(0,1)) + 
  scale_colour_manual(values = c("A_hudsonica" = "#1B9E99",
                               "A_lilljeborgi" = "#D95F02",
                               "F" = "#7570B3",
                               "IV" = "#E7298A",
                               "out_group" = "#666666",
                               "S" = "#66A61E",
                               "SB" = "#E6AB02",
                               "X" = "#A6761D"),
                    na.value = "black") + 
  theme_matt_facets() + 
  theme(legend.position = "none")

ggarrange(lat_clade_prop_plot, temp_clade_prop_plot)
```

Using a Loess smoother (span = 2), highlights how the proportional abundance of the clades is affected by temperature in different ways. Clade F occurs in high proportion at high temperature, Clade IV occurs in high proportions at low temperatures, while Clade X occurs in high proportions at intermediate temperatures. Clade S generally occurs in low proportions, with a few exceptions.  

```{r fig.width=8, fig.height=8}
ggplot(clade_prop, aes(x = collection_temp, y = prop, colour = Clade)) + 
  facet_wrap(Clade~.) + 
  geom_point(size = 3) + 
  geom_smooth(span = 2,
    se = T, 
    linewidth = 3) + 
  labs(x = "Collection Temp. (°C)", 
       y = "Proportion") + 
  coord_cartesian(ylim = c(0,1)) + 
  scale_y_continuous(breaks = c(0,1)) + 
  scale_colour_manual(values = c("A_hudsonica" = "#1B9E99",
                               "A_lilljeborgi" = "#D95F02",
                               "F" = "#7570B3",
                               "IV" = "#E7298A",
                               "out_group" = "#666666",
                               "S" = "#66A61E",
                               "SB" = "#E6AB02",
                               "X" = "#A6761D"),
                    na.value = "black") + 
  theme_matt_facets() + 
  theme(legend.position = "none")
```

### PCA

```{r all-pca, fig.width=8, fig.height=6.5}
tonsa_inds = which(all_samples$Clade != "A_hudsonica")
tonsa_samples = filter(all_samples, Clade != "A_hudsonica")
tonsa_matrix = pcangsd_matrix[tonsa_inds, tonsa_inds]
tonsa_eigs = data.frame(eigen(tonsa_matrix)$vectors[,1:2])
pc1_var = round((eigen(tonsa_matrix)$values[1] / sum(eigen(tonsa_matrix)$values)) * 100, digits = 2)
pc2_var = round((eigen(tonsa_matrix)$values[2] / sum(eigen(tonsa_matrix)$values)) * 100, digits = 2)
tonsa_pca_df = tonsa_eigs %>% 
  bind_cols(tonsa_samples) %>% 
  select(-collection_date, -site, -run, -warming_tol, -doy)

ggplot(tonsa_pca_df, aes(x = X1, y = X2, colour = Clade)) + 
  geom_point(size = 5) + 
  labs(x = paste0("PC1 (", pc1_var, "%)", sep = ""),
       y = paste0("PC2 (", pc2_var, "%)", sep = "")) + 
  scale_colour_manual(values = c("F" = "#b8cfff",
                                 "IV" = "#fc4e91",
                                 "S" = "#4b751c",
                                 "X" = "#ed9e0c")) + 
  theme_matt() + 
  theme(legend.position = "right")

```


```{r clade-pca, fig.width=8, fig.height=6.5}

f_inds = which(all_samples$Clade == "F")
f_samples = filter(all_samples, Clade == "F")
f_matrix = pcangsd_matrix[f_inds, f_inds]
f_eigs = data.frame(eigen(f_matrix)$vectors[,1:2])
f_pc1_var = round((eigen(f_matrix)$values[1] / sum(eigen(f_matrix)$values)) * 100, digits = 2)
f_pc2_var = round((eigen(f_matrix)$values[2] / sum(eigen(f_matrix)$values)) * 100, digits = 2)
f_pca_df = f_eigs %>% 
  bind_cols(f_samples) %>% 
  select(-collection_date, -site, -run, -warming_tol, -doy)

f_pca_df %>% 
  ggplot(aes(x = X1, y = X2, colour = pop)) + 
  #facet_wrap(season~.) + 
  geom_point(size = 3) + 
  labs(x = paste0("PC1 (", f_pc1_var, "%)", sep = ""),
       y = paste0("PC2 (", f_pc2_var, "%)", sep = ""),
       title = "Clade - F") + 
  scale_color_manual(values = pop_cols)  + 
  theme_matt_facets()

#########

x_inds = which(all_samples$Clade == "X")
x_samples = filter(all_samples, Clade == "X")
x_matrix = pcangsd_matrix[x_inds, x_inds]
x_eigs = data.frame(eigen(x_matrix)$vectors[,1:2])
x_pc1_var = round((eigen(x_matrix)$values[1] / sum(eigen(x_matrix)$values)) * 100, digits = 2)
x_pc2_var = round((eigen(x_matrix)$values[2] / sum(eigen(x_matrix)$values)) * 100, digits = 2)
x_pca_df = x_eigs %>% 
  bind_cols(x_samples) %>% 
  select(-collection_date, -site, -run, -warming_tol, -doy)

x_pca_df %>% 
  ggplot(aes(x = X1, y= X2, colour = pop)) + 
  #facet_wrap(season~.) + 
  geom_point(size = 3) + 
  labs(x = paste0("PC1 (", x_pc1_var, "%)", sep = ""),
       y = paste0("PC2 (", x_pc2_var, "%)", sep = ""),
       title = "Clade - X") + 
  scale_color_manual(values = pop_cols)  + 
  theme_matt_facets()

#########

s_inds = which(all_samples$Clade == "S")
s_samples = filter(all_samples, Clade == "S")
s_matrix = pcangsd_matrix[s_inds, s_inds]
s_eigs = data.frame(eigen(s_matrix)$vectors[,1:2])
s_pc1_var = round((eigen(s_matrix)$values[1] / sum(eigen(s_matrix)$values)) * 100, digits = 2)
s_pc2_var = round((eigen(s_matrix)$values[2] / sum(eigen(s_matrix)$values)) * 100, digits = 2)
s_pca_df = s_eigs %>% 
  bind_cols(s_samples) %>% 
  select(-collection_date, -site, -run, -warming_tol, -doy)

s_pca_df %>% 
  ggplot(aes(x = X1,y = X2, colour = pop)) + 
  #facet_wrap(season~.) + 
  geom_point(size = 3) + 
  labs(x = paste0("PC1 (", s_pc1_var, "%)", sep = ""),
       y = paste0("PC2 (", s_pc2_var, "%)", sep = ""),
       title = "Clade - S") + 
  scale_color_manual(values = pop_cols)  + 
  theme_matt_facets()

#########

IV_inds = which(all_samples$Clade == "IV")
IV_samples = filter(all_samples, Clade == "IV")
IV_matrix = pcangsd_matrix[IV_inds, IV_inds]
IV_eigs = data.frame(eigen(IV_matrix)$vectors[,1:2])
IV_pc1_var = round((eigen(IV_matrix)$values[1] / sum(eigen(IV_matrix)$values)) * 100, digits = 2)
IV_pc2_var = round((eigen(IV_matrix)$values[2] / sum(eigen(IV_matrix)$values)) * 100, digits = 2)
IV_pca_df = IV_eigs %>% 
  bind_cols(IV_samples) %>% 
  select(-collection_date, -site, -run, -warming_tol, -doy)

IV_pca_df %>% 
  ggplot(aes(x = X1,y = X2, colour = pop)) + 
  #facet_wrap(season~.) + 
  geom_point(size = 3) + 
  labs(x = paste0("PC1 (", IV_pc1_var, "%)", sep = ""),
       y = paste0("PC2 (", IV_pc2_var, "%)", sep = ""), 
       title = "Clade - IV") + 
  scale_color_manual(values = pop_cols)  + 
  theme_matt_facets()
```



## Misc. Details

```{r temp-record-plot, fig.width=6, fig.height=6}
ggplot(temp_record, aes(x = minute_passed, y = temp_C, group = factor(run))) + 
  geom_abline(slope = 0.3, intercept = mean(temp_record[temp_record$minute_interval == 0, 8])) + 
  geom_abline(slope = 0.1, intercept = mean(temp_record[temp_record$minute_interval == 0, 8])) + 
  geom_line(linewidth = 0.2, alpha = 0.8) + 
  geom_point(data = full_data, 
             aes(x = time, y = ctmax + 0.4),
             size = 2,
             shape = 25) +
  labs(x = "Time passed (minutes)",
       y = "Temperature (degrees C)",
       fill = "Trial Number") + 
  guides(colour = "none") + 
  theme_matt(base_size = 16) + 
  theme(legend.position = "right")
```

```{r ramp-record-plot, fig.width=6, fig.height=6}
ramp_record2 = ramp_record %>% 
  group_by(run, minute_interval) %>% 
  summarise(mean_ramp = mean(ramp_per_minute)) %>% 
  ungroup()

ggplot(ramp_record2, aes(x = minute_interval, y = mean_ramp)) + 
  geom_hline(yintercept = 0.3) + 
  geom_hline(yintercept = 0.1) + 
  #geom_point() + 
  geom_hex(bins = 30) + 
  ylim(0, 0.35) + 
  labs(y = "Ramp Rate (deg. C / min.)",
       x = "Time into run (minute)") + 
  theme_matt(base_size = 16) 
```

```{r rep-comp, fig.height=12, fig.width=8}
full_data %>% 
  drop_na(replicate) %>%  
  ggplot(aes(x = factor(replicate), y = ctmax, group = site)) + 
  facet_grid(site~season, scales = "free_y") + 
  geom_point(position = position_jitter(width = 0.1, height = 0),
             alpha = 0.4,
             colour = "grey30") + 
  geom_smooth(method = "lm", colour = "black") + 
  labs(x = "Replicate", 
       y = "CTmax") + 
  theme_matt_facets()
```

```{r}

early_peak = full_data %>% 
  filter(season %in% c("early", "peak")) %>% 
  mutate(season = if_else(season == "early", "one", "two")) %>% 
  group_by(site) %>% 
  mutate(ctmax_sd_p = sd(ctmax),
         size_sd_p = sd(size), 
         temp_change = max(collection_temp) - min(collection_temp),
         avg_temp = (max(collection_temp) + min(collection_temp)) / 2,
         days_passed = max(doy) - min(doy)) %>% 
  select(site, lat, season, 
         ctmax_sd_p, size_sd_p, 
         temp_change, avg_temp, days_passed, 
         ctmax, size) %>%
  group_by(site, lat, season, 
           ctmax_sd_p, size_sd_p, 
           temp_change, avg_temp, days_passed) %>% 
  summarize(ctmax = mean(ctmax),
            size = mean(size)) %>% 
  pivot_wider(id_cols = c(site, lat, ctmax_sd_p, size_sd_p, 
                          temp_change, avg_temp, days_passed), 
              names_from = season, 
              values_from = c(ctmax, size)) %>% 
  mutate(season = "early_to_peak") %>%  
  drop_na()

peak_late = full_data %>% 
  filter(season %in% c("peak", "late")) %>% 
  mutate(season = if_else(season == "peak", "one", "two")) %>% 
  group_by(site) %>% 
  mutate(ctmax_sd_p = sd(ctmax),
         size_sd_p = sd(size), 
         temp_change = last(collection_temp) - first(collection_temp),
         avg_temp = (max(collection_temp) + min(collection_temp)) / 2,
         days_passed = max(doy) - min(doy)) %>% 
  select(site, lat, season, ctmax_sd_p, size_sd_p, 
         temp_change, avg_temp, days_passed, 
         ctmax, size) %>%
  group_by(site, lat, season, ctmax_sd_p, size_sd_p, 
           temp_change, avg_temp, days_passed) %>% 
  summarize(ctmax = mean(ctmax),
            size = mean(size)) %>% 
  pivot_wider(id_cols = c(site, lat, ctmax_sd_p, size_sd_p, 
                          temp_change, avg_temp, days_passed), 
              names_from = season, 
              values_from = c(ctmax, size)) %>% 
  mutate(season = "peak_to_late") %>%  
  drop_na()
  

generations = bind_rows(early_peak, peak_late) %>% 
  mutate("gen_time" = 5490*(avg_temp + 1)^-2.05,
         "gens" = floor(days_passed / gen_time))
```

```{r num-gens-plot, fig.width=10, fig.height=7}
ggplot(generations, aes(x = lat, y = gens, colour = site, shape = season)) + 
  geom_hline(yintercept = 1) + 
  geom_point(size = 5) + 
  scale_colour_manual(values = site_cols) + 
  labs(x = "Latitude", 
       y = "Generations between \ncollections") +
  scale_y_continuous(breaks = seq(from = 0, to = 21, by = 5)) + 
  theme_matt() + 
  theme(legend.position = "right")
```

```{r rank-sims, fig.width=18, fig.height=8}
obs_ranks = ggplot(full_data, aes(x = rank)) + 
  facet_wrap(tube~.) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(breaks = c(2,4,6,8,10)) + 
  ggtitle("Observation") + 
  theme_matt_facets()

sim_data = data.frame()
for(i in 1:max(full_data$run)){
  rep_data = data.frame("tube" = sample(c(1:10), size = 10, replace = F), 
                        "rank" = c(1:10),
                        "rep" = i) %>% 
    arrange(tube)
  
  sim_data = bind_rows(sim_data, rep_data)
  
}

sim_ranks = ggplot(sim_data, aes(x = rank)) + 
  facet_wrap(tube~.) + 
  geom_histogram(binwidth = 1) + 
  scale_x_continuous(breaks = c(2,4,6,8,10)) + 
  ggtitle("Simulation") + 
  theme_matt_facets()


ggarrange(obs_ranks, sim_ranks)
```

