---
title: "Comparing seasonal and latitudinal patterns in thermal adaptation"
date: "`r Sys.Date()`"
output: 
  html_document:
          code_folding: hide
          code_download: true
          toc: true
          toc_float: true
  github_document:
          html_preview: false
          toc: true
          toc_depth: 3
---

```{r setup, include=T, message = F, warning = F, echo = F}

knitr::opts_chunk$set(
  echo = knitr::is_html_output(),
  fig.align = "center",
  fig.path = "../Figures/markdown/",
  dev = c("png", "pdf"),
  message = FALSE,
  warning = FALSE,
  collapse = T
)

theme_matt = function(base_size = 18,
                      dark_text = "grey20"){
  mid_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[2]
  light_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[3]
  
  ggpubr::theme_pubr(base_family="sans") %+replace% 
    theme(
      panel.background  = element_rect(fill="transparent", colour=NA), 
      plot.background = element_rect(fill="transparent", colour=NA), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      text = element_text(colour = mid_text, lineheight = 1.1),
      title = element_text(size = base_size * 1.5,
                           colour = dark_text),
      axis.text = element_text(size = base_size,
                               colour = mid_text),
      axis.title.x = element_text(size = base_size * 1.2,
                                  margin = unit(c(3, 0, 0, 0), "mm")),
      axis.title.y = element_text(size = base_size * 1.2,
                                  margin = unit(c(0, 5, 0, 0), "mm"), 
                                  angle = 90),
      legend.text = element_text(size=base_size * 0.9),
      legend.title = element_text(size = base_size * 0.9, 
                                  face = "bold"),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25,"cm")
    )
}

theme_matt_facets = function(base_size = 18,
                             dark_text = "grey20"){
  mid_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[2]
  light_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[3]
  
  theme_bw(base_family="sans") %+replace% 
    theme(
      panel.grid = element_blank(),
      panel.background  = element_rect(fill="transparent", colour=NA), 
      plot.background = element_rect(fill="transparent", colour=NA), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      text = element_text(colour = mid_text, lineheight = 1.1),
      strip.text.x = element_text(size = base_size),
      title = element_text(size = base_size * 1.5,
                           colour = dark_text),
      axis.text = element_text(size = base_size,
                               colour = mid_text),
      axis.title.x = element_text(size = base_size * 1.2,
                                  margin = unit(c(3, 0, 0, 0), "mm")),
      axis.title.y = element_text(size = base_size * 1.2,
                                  margin = unit(c(0, 5, 0, 0), "mm"), 
                                  angle = 90),
      legend.text = element_text(size=base_size * 0.9),
      legend.title = element_text(size = base_size * 0.9, 
                                  face = "bold"),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25,"cm")
    )
}

site_cols = c("Key Largo" = "indianred4", 
              "Manatee River" = "coral3", 
              "St. Petersburg" = "coral",
              "Tyler Cove" = "goldenrod1",
              "Ganey's Wharf" = "darkgoldenrod3", 
              "Esker Point" = "darkseagreen3",
              "Sawyer Park" = "palegreen4", 
              "St. Thomas de Kent Wharf" = "deepskyblue3",
              "Miramichi" = "dodgerblue4")
```

## Site Characteristics 
Copepods were collected by surface tow from sites across the Western Atlantic at several times throughout the year. The sites are shown below. Temperatures at the time of collection were measured using a manual thermometer.    

```{r site-chars, fig.width=10, fig.height=6}
coords = site_data %>%
  select(site, long, lat) %>%
  distinct()

site_map = map_data("world") %>% 
  filter(region %in% c("USA", "Canada")) %>% 
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group),
               fill = "lightgrey") + 
  coord_map(xlim = c(-85,-60),
            ylim = c(25, 48)) + 
  geom_point(data = coords,
             mapping = aes(x = long, y = lat, colour = site),
             size = 3) +
  scale_colour_manual(values = site_cols) + 
  labs(x = "Longitude", 
       y = "Latitude") + 
  theme_matt(base_size = 16)

site_temp_plot = full_data %>% 
  select(site, season, doy, collection_temp, collection_salinity) %>%  
  distinct() %>% 
  ggplot(aes(x = doy, y = collection_temp, colour = site)) + 
  geom_point(size = 6) + 
  geom_line(linewidth = 1) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Temperature (°C)",
       x = "Day of the Year") +
  theme_matt() + 
  theme(legend.position = "right")

ggpubr::ggarrange(site_map, site_temp_plot, common.legend = T, legend = "bottom")
```

Exact locations for the sites are provided here. The two sites in Maryland are differentiated by their salinity: Tyler's Cove has a relatively high salinity (15 psu) compared to Ganey's Wharf (2 psu).  

```{r site-tab}
site_data %>%  
  arrange(lat) %>%  
  select("Site" = site, "Region" = region, "Lat" = lat, "Long" = long) %>% 
  knitr::kable(align = "c")
```


## Critical Thermal Limits    
Critical thermal maxima (CTmax) was measured using a custom setup. The method uses a standard dynamic ramping assay to determine the maximum temperature individuals could sustain normal functioning. This differs from lethal temperatures, and indeed, all individuals observed so far recovered following the assay. 

Individuals were rested for one hour after collection before the assay. During the assay, copepods were held in 0.2 um filtered seawater, adjusted to match the salinity at the time of collection with bottled spring water. During the assay, several 'control' individuals were maintained in this adjusted salinity solution, but did not experience the temperature ramp, to ensure that there was no background mortality. 

Shown below are the measured CTmax values. 

```{r seasonal-ct-max}
ggplot(full_data, aes(x = season, y = ctmax, colour = site)) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.5)) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "right")
```

## Body Size    
Following the CTmax assay, individuals were photographed for body size measurements. Prosome lengths were measured from these photographs using a scale micrometer and the software ImageJ. These measurements are shown below. 

```{r seasonal-body-size}
ggplot(full_data, aes(x = season, y = size, colour = site)) + 
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0,
                                             dodge.width = 0.5)) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Season") +
  theme_matt() + 
  theme(legend.position = "right")
```

## Trait Correlations     

We expect that collections from warmer waters should yield copepods with higher thermal limits and smaller body sizes. 

```{r temp-cors, fig.width=10, fig.height=6}
ctmax_temp_plot = ggplot(full_data, aes(x = collection_temp, y = ctmax)) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2, 
              colour = "grey") + 
  geom_point(aes(colour = site), 
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

size_temp_plot = ggplot(full_data, aes(x = collection_temp, y = size)) + 
  geom_smooth(method = "lm", se = T,
              linewidth = 2, 
              colour = "grey") + 
  geom_point(aes(colour = site), 
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "Prosome Length (mm)",
       x = "Collection Temp. (°C)") +
  theme_matt() + 
  theme(legend.position = "right")

ggpubr::ggarrange(ctmax_temp_plot, size_temp_plot, common.legend = T, legend = "bottom")
```


Of particular interest is the relationship between prosome length and CTmax. In many cases, larger body sizes are associated with cold adaptation/acclimation. We may therefore see this pattern emerge across populations or seasons. If populations contain a mix of cold- and warm-adapted genotypes, however, we might also see this relationship emerge **within** individual collections. Shown below is the relationship between prosome length and CTmax for the individuals measured thus far. Individual regression lines for each site are shown along with a 'universal' regression in grey. 

```{r ctmax-vs-size, fig.width=8, fig.height=4}
full_data %>%  
  filter(ctmax > 31) %>% 
  ggplot(aes(x = size, y = ctmax)) + 
  # geom_smooth(data = filter(full_data, ctmax > 31), 
  #             aes(x = size, y = ctmax),
  #             method = "lm", 
  #             colour = "grey60", 
  #             se = F,
  #             linewidth = 2) + 
  geom_smooth(method = "lm", se = F,
              linewidth = 2,
              colour = "grey70") + 
  geom_point(aes(colour = site),
             size = 2, alpha = 0.7) + 
  scale_colour_manual(values = site_cols) + 
  labs(y = "CTmax (°C)",
       x = "Prosome Length (mm)") +
  theme_matt() + 
  theme(legend.position = "right")
```

## Next Steps     

After phenotyping, each individual was preserved in 95% ethanol. Individual DNA libraries will be prepared using Twist Bio 96-plex prep kits, then sequenced on an Illumina NovaSeq X Plus. Using the low-coverage whole genome sequences, we will examine seasonal patterns in allele frequency change, and compare these fine scale temporal patterns with the larger latitudinal patterns in allele frequency to determine whether the same alleles driving rapid seasonal adaptation are in play over larger spatial (and longer temporal) scales. 

## Misc. Details    

```{r temp-record-plot, fig.width=6, fig.height=6}
ggplot(temp_record, aes(x = minute_passed, y = temp_C, group = factor(run))) + 
  geom_abline(slope = 0.3, intercept = mean(temp_record[temp_record$minute_interval == 0, 8])) + 
  geom_abline(slope = 0.1, intercept = mean(temp_record[temp_record$minute_interval == 0, 8])) + 
  geom_line(linewidth = 0.2, alpha = 0.8) + 
  geom_point(data = full_data, 
             aes(x = time, y = ctmax + 0.4),
             size = 2,
             shape = 25) +
  labs(x = "Time passed (minutes)",
       y = "Temperature (degrees C)",
       fill = "Trial Number") + 
  guides(colour = "none") + 
  theme_matt(base_size = 16) + 
  theme(legend.position = "right")
```

```{r ramp-record-plot, fig.width=6, fig.height=6}
ramp_record2 = ramp_record %>% 
  group_by(run, minute_interval) %>% 
  summarise(mean_ramp = mean(ramp_per_minute)) %>% 
  ungroup()

ggplot(ramp_record2, aes(x = minute_interval, y = mean_ramp)) + 
  geom_hline(yintercept = 0.3) + 
  geom_hline(yintercept = 0.1) + 
  #geom_point() + 
  geom_hex(bins = 30) + 
  ylim(0, 0.35) + 
  labs(y = "Ramp Rate (deg. C / min.)",
       x = "Time into run (minute)") + 
  theme_matt(base_size = 16) 
```

